Mechanisms of detachment in fibrillar adhesive systems

### 2022-08-13 source/tex/manuscript.tex

\journal{Acta Biomaterialia}
\usepackage[T2A,LGR,T1]{fontenc}
%\usepackage{babel}
\usepackage{textcomp}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stackrel}
\usepackage{graphicx}
\usepackage{subscript}
%\usepackage[unicode=true,pdfusetitle,
% bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
% breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
% {hyperref}

@@

\begin{abstract}
Several creatures can climb on smooth surfaces with the help of hairy
adhesive pads on their legs. A rapid change from strong attachment
to effortless detachment of the leg is enabled by the asymmetric geometry
of the tarsal hairs. In this study, we propose mechanisms by which
a hairy pad can be easily detached, even when the hairs possess no
asymmetry. Here, we examine the possible function of the tibia-tarsus (t-t) leg joint and the claws.
Based on a spring-based model, we consider three modes
of detachment: vertically pulling the pad while maintaining either
a 1) fixed or a 2) free t-t leg joint, or by 3) flexing the pad about
the claw. We show that in all cases, the adhesion force can be significantly
reduced due to elastic forces when the hairs deform non-uniformly
across the array. %, an effect that we call ``\emph{elastic weakening}''.
Our proposed model illustrates the design advantage provided by such fibrillar
adhesive systems, which not only provides strong adhesion, but also
allows easy detachment, making it suitable as organs for fast locomotion.
The presented approaches can be potentially used to switch the adhesion
state in bio-inspired fibrillar adhesives, by incorporating artificial joints and claws into their design, 
without the need of asymmetric or stimuli-responsive fibrillar structures.
\end{abstract}

@@

Over the past few decades, there have been numerous studies to understand
how animals such as geckos and insects are able to walk on surfaces
of any direction while defying gravity. A microscopic observation
reveals that, in many cases, animals have a dense array of fibrillar structures
at the end of their legs \cite{RN198,RN59}. These "hairy" adhesive
pads help the animal to stay attached to any surface or detach easily
at will for countless cycles, a property that is referred to as \emph{reversible
adhesion}. Previous attempts to theoretically explain adhesion in
hairy pads \cite{RN40,RN120} has followed two fundamental approaches:
either by energy balance, or by force balance.

In the energy balance approach, adhesion is usually characterized by ``work
of adhesion'' (W\textsubscript{adh}), which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, which increases W\textsubscript{adh}
and thus adhesion is enhanced \cite{RN29,RN114}. Detachment of an
individual hair can be explained based on Kendall's peeling theory \cite{RN147, RN296},
which predicts low adhesion at high peeling angles.

In the force balance approach, adhesion is characterized by pull-off
force, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a ``cohesive zone model'', Hui et. al. \cite{RN72}
had identified two regimes of single hair detachment: 1) a \emph{flaw sensitive}
regime, where, for large hair radius, contact failure occurs due to crack propagation, 
initiated by a stress singularity at the edge of the hair, leading
to low \textgreek{sv}\textsubscript{p}; 2) a \emph{flaw insensitive} regime,
where, for small hair radius, the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Tian et. al. \cite{RN230} had shown
that the ``spatula'' shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. Federle \cite{RN20}
had further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally, and easily detached by elastic
recoil when pushed distally.

The theories presented so far suggests that low detachment force of
a fibrillar adhesive pad can be attained by 1) increasing the stress concentration
by peeling the pad at high angles, or 2) laterally shearing the pad
before pull-off, which requires the hairs to have an asymmetric geometry
or curvature. However, some insects like male dock beetles predominantly
possess flat ``mushroom'' tipped hairs on their pads \cite{RN19},
which are relatively less asymmetric compared to the ``spatula''
tipped hairs. This begs the question of how the beetle can still easily
detach their legs during locomotion, since these mushroom shaped hairs
should be resilient to detachment via lateral shearing. Artificial
micro-pillar adhesives mimicking the mushroom shaped hairs of the
beetle have in fact demonstrated its superior adhesive strength in
comparison to other geometries \cite{RN112,RN57}. Introducing asymmetry
into the pillar geometry of such mimics for easy detachment is however
challenging due to current limitations in fabrication techniques and
difficulty in scaling-up for practical applications \cite{RN57}.
Alternate strategies are thus desired to switch the adhesion state
of symmetric pillar arrays in a reversible manner, e.g. by buckling
the pillars under compressive load leading to their contact loss \cite{RN279}
or by using special materials reacting to external stimuli such as
magnetic field \cite{RN186}, UV light \cite{RN280} or temperature \cite{RN281}.

In this work, we propose the possible mechanisms by which adhesive
pads with symmetric hairs can be easily detached under a theoretical
framework. The hairy pad is modelled as an array of springs and the
force-balance approach is used to quantify the net adhesion \cite{RN117, RN253}.
Here, we focus our analysis to only normal adhesion force necessary to separate 
the pad from a flat surface.
We hope our work to provide strategies to control the adhesion force
of an artificial micro-pillar adhesive, which has applications in
bio-inspired climbing robots, pick-and-place operations and reusable
adhesives.

@@

The fibrillar adhesive pad is assumed to be a one dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg, assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge, analogous to the tarsus-tibia (t-t) leg joint of an insect, 
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an instantaneous angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net force on the pad and $M_{net}$ is the the net
moment about the hinge, at a particular instant during the detachment
process. The contact points between the hair and the surface is assumed to be laterally fixed due to high friction forces.

\begin{figure}[H]
\includegraphics{Figure1-model_schematic}\caption{\textbf{Spring contact model of a fibrillar adhesive pad.} The pad consists of
an array of $N_{t}$ hairs connected to a deformable leg at the hinge.
At a particular distance, $d_{s}$, $n$ number of hairs are in contact
and the array is oriented at a tilt angle, $\theta$, with the surface.
\label{fig:Schematic}}
\end{figure}

@@

where, $\Psi=s-\frac{n-1}{2}w$. For a particular value of $n$, equation
\ref{eq:force_n_d} is valid until a certain distance, $d_{s,max}$,
above which the left most hair will detach. Just before detachment,
this hair will be at its maximum length, $l_{h,p}$. From simple geometry
we can thus find:

@@

Equation \ref{eq:force_n_d} will be valid for $d_{s}\leq d_{s,max}$.

@@

We consider three possible mechanisms to detach the adhesive pad from
a surface: 1) \emph{Fixed pull}, where the pad is pulled vertically
up while keeping a fixed hinge (t-t leg joint); 2) \emph{Free pull,} where the
pad is pulled vertically up while keeping the hinge (t-t leg joint) free to allow
rotation of the array; 3) \emph{Flex}, where the pad is hinged
to an external point (claw-hinge), and detached in a rotary fashion,
emulating the claw function in insects. To investigate each case in
detail, let us assume a pad to be a one dimensional analogue of a 
dock beetle's adhesive pad \cite{RN19, RN79} with $N_{t}=25$ hairs and $\hat{f_{p}}=0.1$
attached to a stiff leg (tibia) with $k_{l}\rightarrow\infty$. The situation
of a soft leg (tibia) with $k_{l}/k_{h}=10$ is also considered for the first
two cases involving vertical detachment.

@@

\begin{figure}[H]
\includegraphics{Case1-Fixed_pull}\caption{\textbf{Detachment by \emph{Fixed Pull}.} 
Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a fixed hinge (tarsus-tibia leg joint). The
tilt angle, $\theta$, of the array is kept fixed during detachment.
The leg (tibia) is either stiff ($k_{l}\rightarrow\infty$) or soft ($k_{l}/k_{h}=10$).
Positive force values represent attraction. The green dashed line
represents the maximum possible adhesion for the pad. All values are
normalized to dimensionless forms, as described in text.\label{fig:Fixed-pull}}
\end{figure}

The fibrillar adhesive pad can be detached by pulling it vertically upwards
while maintaining a constant tilt angle, $\theta$, with the surface.
This can be achieved if the hinge (t-t leg joint) is kept fixed. 
Equations \ref{eq:force_n_d}, \ref{eq:d_max}
and \ref{eq:simul_detach} can be used to get the resulting force-distance
curves.

Increasing the tilt of the pad decreases its maximum force or adhesion
(Figure \ref{fig:Fixed-pull}). Tilting the pad causes an inhomogeneous
deformation of hairs, where, on one end they are stretched, while,
on the other end they are compressed. The balance of the respective
attractive and repulsive elastic forces of the hairs ultimately results
in a decrease in the net force. The tilted orientation also causes
the individual hairs to detach distinctly rather than simultaneously,
further reducing the net adhesive power of the array. We term this
effect of loss in adhesion due to a non-uniform hair deformation across
the array as ``\emph{elastic weakening}''. When there is no tilt
($\theta=0\lyxmathsym{\textdegree}$), all the hairs undergo identical
deformation and ultimately detaches simultaneously after a distance,
$\hat{d_{s}}=0.1$. Here, no \emph{elastic weakening} occurs and the
pad shows the maximum possible adhesion.

For the case of a stiff leg (tibia), we see that at small distances, all hairs
of the pad are in contact with the surface, resulting in a linear
force response. On further pulling, the hairs will start to detach
sequentially from left to right, indicated by a characteristic ``saw-tooth''
jitter in the force curves. The hairs of the pad with a higher tilt
angle will start to detach first, followed by the ones with a lower
tilt.

For the case of a soft leg (tibia), we observe a similar effect of tilt angle
on the force curves as before. The maximum adhesion force at a particular
tilt is the same as that for the stiff leg. The saw-tooth jitter are
however minimized due to the leg's deformation, leading to a ``dampened''
force response. Interestingly, the force abruptly drops to zero for
the angles $0.2\lyxmathsym{\textdegree}$ and $0.4\lyxmathsym{\textdegree}$.
This is an effect of the elastic recoil of the leg while each hair
loses contact during the detachment process (equation \ref{eq:backing_deform}).
The length difference between the detached hair just before it breaks
contact and its adjacent hair is $w\sin\theta$. If the leg's recoil
length, $\Delta l>w\sin\theta$, the adjacent hair will be stretched
more than its maximum length ($l_{h,p}$), and thus will also detach,
leading to further recoil of the leg. Equation \ref{eq:backing_deform}
shows that $\Delta l$ increases with every subsequent loss of hair
contact if $\theta$ is kept constant. This implies that, once initiated,
the leg's recoil will always be large enough to detach every remaining
hair, resulting in a spontaneous propagation of the detachment front
until the pad completely breaks contact with the surface. This is consistent with 
a recent report of catastrophic failure, due to a similar recoil effect of the measurement system,
seen in micro-fibrillar adhesives with a narrow variance of individual fibril adhesive strengths \cite{RN273}.

@@

\includegraphics{Case2-Free_pull}\caption{\textbf{Detachment by \emph{Free Pull}.} 
Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a free hinge (tarsus-tibia leg joint). $s$
is the distance between the leg joint and the right end of the array
and $L$ is the array length. The free joint allows further tilting
of the array during the vertical pull. The leg (tibia) is either stiff ($k_{l}\rightarrow\infty$)
or soft ($k_{l}/k_{h}=10$). Positive force values represent attraction.
The green dashed line represents the maximum possible adhesion for
the pad. All values are normalized to dimensionless forms, as described
in text. \label{fig:Free-pull}}

@@

Similar to the previous case, we once again consider the situation
where the adhesive pad is pulled vertically upwards for detachment.
However now, the leg joint is assumed to be free. In this case, the
array will reorient itself to maintain a zero net torque about the
joint during the entire detachment process. At any given instant,
the tilt angle, $\theta$, can be found by setting $M_{net}$ to zero
in equation \ref{eq:moment_n_d} to get:

@@

Using the above relation together with equations \ref{eq:force_n_d},
\ref{eq:d_max} and \ref{eq:simul_detach}, we can find force-distance
curves during a free vertical pull of the adhesive pad. Since the
position of the leg joint will influence the net torque, we use the
ratio, $s/L$, to study its effect on the detachment forces.

Maximum adhesion is seen when the leg joint is positioned at the centre
of the array, i.e. $s/L=0.5$ (Figure \ref{fig:Free-pull}). Here,
the net torque due to the hairs is balanced by symmetry and the array
remains parallel to the substrate until all hairs detach simultaneously
at $\hat{d_{s}}=0.1$. Shifting the position of the joint further
away from the array centre leads to lower forces or adhesion. The
resulting torque imbalance will tilt the array, which reduces the
net force due to the \emph{elastic weakening} effect, as described
in the previous section. Higher $s/L$ increases the net torque to
be balanced, leading to a higher tilt of the array and thus lower
net force.

The force curves look qualitatively different compared to the previous
case of \emph{fixed pull}. A sharp maxima is seen, coinciding with
the point when the first hair detaches. Beyond this, the force starts
to decrease sharply and once again shows the characteristic saw-tooth
jitter as the subsequent hairs detach in sequence. Nearly identical
trend is seen for both a stiff and a soft leg. The elastic recoil
of the leg does slightly reduce the amplitude of the jitter for the
soft leg case. However, no abrupt drop in the force is seen like before.
As the hairs detach, the array gets tilted more and more, making it
less likely for the recoil length, $\Delta l$, to exceed $w\sin\theta$
and detach the next adjacent hair. Thus here, we don't see any propagation
of the detachment front when the leg is soft.

@@

\begin{figure}[H]
\includegraphics{Case3-Fixed_flex}\caption{\textbf{Detachment by flex.}
Force curves for a fibrillar adhesive pad detached by flexing it about the claw.
$\hat{F}_{pull}=\frac{F_{pull}}{k_{h}w}$ is the normalised pulling
force necessary to apply the torque about the claw-hinge for detachment,
$\hat{F}_{hinge}=\frac{F_{hinge}}{k_{h}w}$ is the normalised reaction
force on the claw-hinge, $\hat{d_{s}}=\frac{d_{s}-l_{h,0}}{w}$ is
the normalized vertical distance of the claw-hinge from the surface.
Here, $s_{l}/L=1$ and $s_{h}/w=10$. The green dashed line represents
the maximum possible adhesion for the pad. \label{fig:Flex}}
\end{figure}

Instead of a vertical pull, the adhesive pad can also be detached
by rotating it about the claw-hinge, located outside the array. Such
a mode of detachment will be driven by a torque applied by the leg
to rotate the pad around the claw-hinge until all the hairs lose contact.
Let $s_{h}$ be the distance between the claw-hinge and the right
end of the array; $s_{l}$ be the distance between the fixed leg joint
and the right end of the array. To illustrate the mechanism, let us
fix $s_{l}/L=1$ and $s_{h}/w=10$ and vary the vertical claw-hinge
distance, $d_{s}$. At any particular instant, the pulling force applied
by the leg, $F_{pull}=M_{net}/\left(s_{l}+s_{h}\right)$, where $M_{net}$
is obtained by setting $s=-s_{h}$ in equation \ref{eq:moment_n_d}.
Equation \ref{eq:force_n_d} will give us the reaction force acting
on the claw-hinge, $F_{claw-hinge}$.

Decreasing the vertical claw-hinge distance reduces the pulling force
necessary to undergo detachment by flexing (Figure \ref{fig:Flex}).
One can imagine that initially, when the array is parallel to the
surface, a lower value of $d_{s}$ means the hairs are in a more compressed
state. When the pad is subsequently rotated around the claw-hinge,
the tilted array will once again lead to an \emph{elastic weakening}
effect due to the inhomogeneous deformation of hairs. This results
in a decrease in the net torque and thus lower $F_{pull}$ for smaller
values of $d_{s}$. $F_{pull}$ can be further reduced of course by
increasing the lever arm ($s_{l}+s_{h}$).

@@

Here, $F_{adh}$ is the adhesion force required to detach the pad
from the surface following a given mechanism and $N_{t}f_{p}$ is
the maximum possible adhesion of the pad (equation \ref{eq:simul_detach}).
Reduction factor, $r$, represents the extent to which the adhesion
can be reduced by choosing the mode of detachment. A large value of
$r$ implies that adhesion can be reduced by a greater factor, and
this mode is more suitable to easily detach.

@@

For the case of detachment via \emph{free pull}, $\hat{f}_{p}$ has
no influence on the reduction factor. On the other hand, shifting
the position of the leg joint further away from the array (i.e. high
$s/L$) results in large values of $r$. In this scenario, the higher
torque exerted by the array leads to a higher tilt, and thus increases
the reduction factor via \emph{elastic weakening}, saturating to the
maximum value of 25.

>>>>

\journal{Journal of Theoretical Biology}
\usepackage[T2A,LGR,T1]{fontenc}
%\usepackage{babel}
\usepackage{textcomp}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stackrel}
\usepackage{graphicx}
\usepackage{subscript}
%\usepackage[unicode=true,pdfusetitle,
% bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
% breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
% {hyperref}

@@

\begin{abstract}
Several creatures can climb on smooth surfaces with the help of hairy
adhesive pads on their legs. A rapid change from strong attachment
to effortless detachment of the leg is enabled by the asymmetric geometry
of the tarsal hairs. In this study, we propose mechanisms by which
the hairy pad can be easily detached, even when the hairs possess no
asymmetry. Here, we examine the possible function of the tibia-tarsus leg joint and the claws. Based on a spring-based model, we consider three modes
of detachment: vertically pulling the pad while maintaining either
a 1) fixed or a 2) free joint, or by 3) flexing the pad about
the claw. We show that in all cases, the adhesion force can be significantly
reduced due to elastic forces when the hairs deform non-uniformly
across the array. %, an effect that we call ``\emph{elastic weakening}''.
Our proposed model illustrates the design advantage of such fibrillar
adhesive systems, that not only provide strong adhesion, but also
allow easy detachment, making them suitable as organs for fast locomotion and reliable hold.
The presented approaches can be potentially used to switch the adhesion
state in bio-inspired fibrillar adhesives, by incorporating artificial joints and claws into their design, without the need of asymmetric or stimuli-responsive fibrillar structures.
\end{abstract}

@@

Over the past few decades, there have been numerous studies to understand
how animals, such as geckos and insects, are able to walk on surfaces
of any direction while seemingly defying gravity. A microscopic observation
reveals that, in many cases, animals have a dense array of fibrillar structures
at the end of their legs \cite{RN198,RN59}. These \emph{hairy} adhesive
pads help the animal to stay attached to any surface or detach easily
at will for countless cycles, a property that is referred to as \emph{reversible
adhesion}. Previous attempts to theoretically explain adhesion in
hairy pads \cite{RN40,RN120,Popov2016BiologicalMW} has followed two fundamental approaches: either by energy balance, or by force balance.

In the energy balance approach, adhesion is usually characterized by \emph{work
of adhesion} (W\textsubscript{adh}), which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, that increases W\textsubscript{adh}
and thus adhesion is enhanced \cite{RN29,RN114}. Detachment of an
individual hair can be explained based on Kendall's peeling theory \cite{RN147, RN296},
which predicts low adhesion at high peeling angles.

In the force balance approach, adhesion is characterized by \emph{pull-off
force}, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a \emph{cohesive zone model}, Hui et. al. \cite{RN72}
have identified two regimes of single hair detachment: 1) a \emph{flaw sensitive}
regime, where, for large hair radius, contact failure occurs due to crack propagation, 
initiated by a stress singularity at the edge of the hair, leading
to low \textgreek{sv}\textsubscript{p}; 2) a \emph{flaw insensitive} regime,
where, for small hair radius, the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Likewise, Tian et. al. \cite{RN230} have shown
that the spatula-shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. The detailed mechanics of the
spatula-shaped hair design for controlling adhesion have been extensively studied by theoretical modelling \cite{pantano2011, sauer2013, Wu2015, kligerman2020, Grill2021, Gouravaraju2021}, artificial mimics \cite{RN57, Murphy2007, Menguc2012, Chary2013, Kim2017} as well as in several biological systems \cite{autumn2000,Langer2004,Varenberg2010}. Federle \cite{RN20}
has further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally and easily detached by elastic
recoil when pushed distally.

The theory presented so far suggests that a low detachment force of
a fibrillar adhesive pad can be achieved either by increasing the stress concentration
by peeling the pad at high angles, or by laterally shearing the pad
before pull-off, which requires the hairs to have an asymmetric geometry
or curvature. However, some insects like male dock beetles predominantly
possess mushroom-shaped hairs with flat discoid terminals on their pads \cite{RN19},
that are relatively less asymmetric compared to the spatula-shaped hairs. These mushroom-shaped hairs have in fact been shown to possess superior
adhesion compared to the spatula-shaped hairs \cite{RN112,RN57} and are generally resistant to detachment via lateral shear \cite{Heepe2014}. Yet, how does the beetles possessing such mushroom-shaped hairs still easily detach their legs during locomotion? Besides, from an application perspective, introducing asymmetry
into the pillar geometry to construct spatula-shaped artificial biomimetic 
adhesives for easy detachment is challenging due to current limitations in fabrication techniques and difficulty in scaling-up \cite{RN57}.
Alternate strategies are thus desired to switch the adhesion state
of symmetric pillar arrays in a reversible manner. This can be achieved, for example, by buckling
the pillars under compressive load leading to their contact loss \cite{RN279}
or by using special materials reacting to external stimuli such as
magnetic field \cite{RN186}, UV light \cite{RN280} or temperature \cite{RN281}.

Employing the force-balance approach, in this work we theoretically model the possible mechanisms by which adhesive pads with axially-symmetric hairs can be easily detached, without the need of any spatula-like asymmetry. 
Here, we focus our analysis on normal adhesion force necessary to separate 
the pad from a flat surface under a purely mechanical action. We found that the maximum force necessary to detach the leg can be significantly reduced by strategically controlling the tilt, joints and claws of the adhesive system. We hope our work to provide new approaches to control the adhesion force
of an artificial micro-pillar adhesive, that has applications in
bio-inspired climbing robots, pick-and-place operations and reusable
adhesives.

@@

Similar to previous approaches \cite{RN117, RN253}, the fibrillar adhesive pad is assumed to be a one-dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg (tibia), assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge, analogous to the tibia-tarsus leg joint of an insect, 
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an instantaneous angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net normal force on the pad and $M_{net}$ is the the net
moment about the hinge, at a particular instant during the detachment
process. We focus only on vertical detachment modes and thus lateral friction forces between the hairs and surface are not considered for our analysis.

\begin{figure}[H]
\includegraphics{Figure1-model_schematic}\caption{\textbf{Spring contact model of a fibrillar adhesive pad.} The pad consists of
an array of $N_{t}$ hairs connected to a deformable leg at the joint.
At a particular distance, $d_{s}$, $n$ number of hairs are in contact
and the array is oriented at a tilt angle, $\theta$, with the surface.
\label{fig:Schematic}}
\end{figure}

@@

where, $\Psi=s-\frac{n-1}{2}w$. For a particular value of $n$, equation
(\ref{eq:force_n_d}) is valid until a certain distance, $d_{s,max}$,
above which the left most hair will detach. Just before detachment,
this hair will be at its maximum length, $l_{h,p}$. From simple geometry
we can thus find:

@@

Equation (\ref{eq:force_n_d}) will be valid for $d_{s}\leq d_{s,max}$.

@@

We consider three tentative scenarios to detach the adhesive pad from
a surface: 1) \emph{Fixed pull}, where the pad is pulled vertically
up while keeping a fixed joint; 2) \emph{Free pull,} where the
pad is pulled vertically up while keeping the joint free to allow
rotation of the array; 3) \emph{Flex}, where the pad is hinged
to an external point (claw-hinge), and detached in a rotary fashion,
emulating the claw function in insects. To investigate each case in
detail, let us assume a pad to be a one-dimensional analogue of a 
dock beetle's adhesive pad \cite{RN19, RN79} with $N_{t}=25$ hairs and $\hat{f_{p}}=0.1$
(see discussion on \emph{detachment pathways} for details) attached to a stiff leg (or tibia) with $k_{l}\rightarrow\infty$. The situation
of a soft leg with $k_{l}/k_{h}=10$ is also considered for the first
two cases involving vertical detachment.

@@

\begin{figure}[H]
\includegraphics{Case1-Fixed_pull}\caption{\textbf{Detachment by \emph{Fixed Pull}.} 
Force-distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a fixed joint. The
tilt angle, $\theta$, of the array is kept fixed during detachment.
The leg (tibia) is either stiff ($k_{l}\rightarrow\infty$) or soft ($k_{l}/k_{h}=10$).
Positive force values represent attraction between the array and the surface. The green dashed line
represents the maximum possible adhesion for the pad. All values are
normalized to dimensionless forms, as described in text.\label{fig:Fixed-pull}}
\end{figure}

The fibrillar adhesive pad can be detached by pulling it vertically upwards
while maintaining a constant tilt angle, $\theta$, with the surface.
This can be achieved if the joint is kept fixed. 
Equations \ref{eq:force_n_d}, \ref{eq:d_max}
and \ref{eq:simul_detach} can be used to get the resulting force-distance
curves for such a scenario.

Increasing the tilt of the pad decreases its maximum force or adhesion
(Figure \ref{fig:Fixed-pull}). Tilting the pad causes an inhomogeneous
deformation of hairs, where, on one end they are stretched, while,
on the other end they are compressed. The balance of the respective
attractive and repulsive elastic forces of the hairs ultimately results
in a decrease in the net force. The tilted orientation also causes
the individual hairs to detach distinctly rather than simultaneously,
further reducing the maximum adhesion of the array. We term this
effect of loss in adhesion due to a non-uniform hair deformation across
the array as \emph{elastic weakening}. When there is no tilt
($\theta=0\lyxmathsym{\textdegree}$), all the hairs undergo identical
deformation and ultimately detach simultaneously after a distance,
$\hat{d_{s}}=0.1$. Here, no \emph{elastic weakening} occurs and the
pad shows the maximum possible adhesion.

For the case of a stiff leg (tibia), we see that at small distances, all hairs
of the pad are in contact with the surface, resulting in a linear
force response. On further pulling, the hairs will start to detach
sequentially from left to right, indicated by a characteristic saw-tooth
jitter in the force curves. The hairs of the pad with a higher tilt
angle will start to detach first, followed by the ones with a lower
tilt.

For the case of a soft leg (tibia), we observe a similar effect of tilt angle
on the force curves as before. The maximum adhesion force at a particular
tilt is the same as that for the stiff leg. The saw-tooth jitter are
however minimized due to the leg's deformation, leading to a dampened
force response. Interestingly, the force abruptly drops to zero for
the angles $0.2\lyxmathsym{\textdegree}$ and $0.4\lyxmathsym{\textdegree}$.
This is an effect of the elastic recoil of the leg while each hair
loses contact during the detachment process (equation (\ref{eq:backing_deform})).
The length difference between the detached hair just before it breaks
contact and its adjacent hair is $w\sin\theta$. If the leg's recoil
length, $\Delta l>w\sin\theta$, the adjacent hair will be stretched
more than its maximum length ($l_{h,p}$), and thus will also detach,
leading to further recoil of the leg. Equation (\ref{eq:backing_deform})
shows that $\Delta l$ increases with every subsequent loss of hair
contact if $\theta$ is kept constant. This implies that, once initiated,
the leg's recoil will always be large enough to detach every remaining
hair, resulting in a spontaneous propagation of the detachment front
until the pad completely breaks contact with the surface. This is consistent with 
a recent report of catastrophic failure, due to a similar recoil effect of the measurement system,
seen in micro-fibrillar adhesives with a narrow variance of individual fibril adhesive strengths \cite{RN273}.

@@

\includegraphics{Case2-Free_pull}\caption{\textbf{Detachment by \emph{Free Pull}.} 
Force-distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a free joint. $s$
is the distance between the joint and the right end of the array
and $L$ is the array length. The free joint allows further tilting
of the array during the vertical pull. The leg (tibia) is either stiff ($k_{l}\rightarrow\infty$)
or soft ($k_{l}/k_{h}=10$). Positive force values represent attraction.
The green dashed line represents the maximum possible adhesion for
the pad. All values are normalized to dimensionless forms, as described
in text. \label{fig:Free-pull}}

@@

Similar to the previous case, we once again consider the situation
where the adhesive pad is pulled vertically upwards for detachment.
However now, the joint is assumed to be freely movable. In this case, the
array will reorient itself to maintain a zero net moment about the
joint during the entire detachment process. At any given instant,
the tilt angle, $\theta$, can be found by setting $M_{net}$ to zero
in equation (\ref{eq:moment_n_d}) to get:

@@

Using the above relation together with equations \ref{eq:force_n_d},
\ref{eq:d_max} and \ref{eq:simul_detach}, we can find force-distance
curves during a free vertical pull of the adhesive pad. Since the
position of the joint will influence the net moment, we use the
ratio, $s/L$, to study its effect on the detachment forces.

Maximum adhesion is seen when the joint is positioned at the centre
of the array, i.e. $s/L=0.5$ (Figure \ref{fig:Free-pull}). Here,
the net moment due to the hairs is balanced by symmetry and the array
remains parallel to the substrate until all hairs detach simultaneously
at $\hat{d_{s}}=0.1$. Shifting the position of the joint further
away from the array centre leads to lower forces or adhesion. The
resulting moment imbalance will tilt the array, which reduces the
net force due to the \emph{elastic weakening} effect, as described
in the previous section. Higher $s/L$ increases the net moment to
be balanced, leading to a higher tilt of the array and thus lower
net force.

The force curves look qualitatively different compared to the previous
case of \emph{fixed pull}. A sharp maxima is seen, coinciding with
the point when the first hair detaches. Beyond this, the force starts
to decrease sharply and once again shows the characteristic saw-tooth
jitter as the subsequent hairs detach in sequence. Nearly identical
trend is seen for both a stiff and a soft leg (tibia). The elastic recoil
of the leg does slightly reduce the amplitude of the jitter for the
soft leg case. However, no abrupt drop in the force is seen like before.
As the hairs detach, the array gets tilted more and more, making it
less likely for the recoil length, $\Delta l$, to exceed $w\sin\theta$
and detach the next adjacent hair. Thus here, we don't see any propagation
of the detachment front when the leg is soft.

@@

\begin{figure}[H]
\includegraphics{Case3-Fixed_flex}\caption{\textbf{Detachment by \emph{Flex}.}
Force curves for a fibrillar adhesive pad detached by flexing it about the claw.
$\hat{F}_{pull}=\frac{F_{pull}}{k_{h}w}$ is the normalised pulling
force necessary to apply the moment about the claw-hinge for detachment,
$\hat{F}_{hinge}=\frac{F_{hinge}}{k_{h}w}$ is the normalised reaction
force on the claw-hinge, $\hat{d_{s}}=\frac{d_{s}-l_{h,0}}{w}$ is
the normalized vertical distance of the claw-hinge from the surface.
Here, $s_{l}/L=1$ and $s_{h}/w=10$. The green dashed line represents
the maximum possible adhesion for the pad. \label{fig:Flex}}
\end{figure}

Instead of a vertical pull, the adhesive pad can also be detached
by rotating it about the claw-hinge, located outside the array. Such
a mode of detachment will be driven by a moment applied by the leg (tibia)
to rotate the pad around the claw-hinge until all the hairs lose contact.
Let $s_{h}$ be the distance between the claw-hinge and the right
end of the array; $s_{l}$ be the distance between the joint
and the right end of the array. The joint is assumed to be fixed here. To illustrate the mechanism, let us
fix $s_{l}/L=1$ and $s_{h}/w=10$ and vary the vertical claw-hinge
distance, $d_{s}$. At any particular instant, the pulling force applied
by the leg, $F_{pull}=M_{net}/\left(s_{l}+s_{h}\right)$, where $M_{net}$
is obtained by setting $s=-s_{h}$ in equation (\ref{eq:moment_n_d}).
Equation (\ref{eq:force_n_d}) will give us the reaction force acting
on the claw-hinge, $F_{claw-hinge}$.

Decreasing the vertical claw-hinge distance reduces the pulling force
necessary to undergo detachment by flexing (Figure \ref{fig:Flex}).
One can imagine that initially, when the array is parallel to the
surface, a lower value of $d_{s}$ means the hairs are in a more compressed
state. When the pad is subsequently rotated around the claw-hinge,
the tilted array will once again lead to an \emph{elastic weakening}
effect due to the inhomogeneous deformation of hairs. This results
in a decrease in the net moment and thus lower $F_{pull}$ for smaller
values of $d_{s}$. $F_{pull}$ can be further reduced of course by
increasing the lever arm ($s_{l}+s_{h}$).

@@

Here, $F_{adh}$ is the adhesion force required to detach the pad
from the surface following a given mechanism and $N_{t}f_{p}$ is
the maximum possible adhesion of the pad (equation (\ref{eq:simul_detach})).
Reduction factor, $r$, represents the extent to which the adhesion
can be reduced by choosing the mode of detachment. A large value of
$r$ implies that adhesion can be reduced by a greater factor, and
this mode is more suitable to easily detach.

@@

For the case of detachment via \emph{free pull}, $\hat{f}_{p}$ has
no influence on the reduction factor. On the other hand, shifting
the position of the joint further away from the array (i.e. high
$s/L$) results in large values of $r$. In this scenario, the higher
moment exerted by the array leads to a higher tilt, and thus increases
the reduction factor via \emph{elastic weakening}, saturating to the
maximum value of 25.
